Multi-Source Debiased Calibration Simulation with Estimated Alpha

### 2024-11-12 src/explore/20241025-mcbssim.qmd
Finish all of the simulations. Added simulation script.

library(dplyr)
library(readr)
library(haven)
library(stringr)
library(purrr)
library(nleqslv)

@@

```{r}
#| label: acs uncertainty
acs_cl <- read_csv("../build/cleaned_acs_data.csv")
acs_med <- filter(acs_cl, HINS3 == 1)

@@

@@

```{r}
#| label: add CMS data
cov_ests <- 
  bind_rows(
  purrr::map_dfr(param_df$varname, function(vn) {
    out <- get_est_uncertainty(acs_med, vn)
    return(tibble(Data = "ACS", Name = vn, Est = out$est, Var = out$var))
  }),
  purrr::map_dfr(param_df$varname[-1], function(vn) {
    out <- get_est_uncertainty(mcbs_all, vn, def_w = "PUFSWGT")
    return(tibble(Data = "MCBS", Name = vn, Est = out$est, Var = out$var))
  }))

@@

cov_ests <- cov_ests %>% 
  add_row(Data = "CMS", Name = "tot", Est = 63892626, Var = 1) %>%
  add_row(Data = "CMS", Name = "sexm", Est = 29159084, Var = 1) %>%
  add_row(Data = "CMS", Name = "sexf", Est = 34733542, Var = 1) %>%
  mutate(Name = ifelse(Name %in% c("HINS3", "ones"), "tot", Name)) %>%
  add_row(Data = "MCBS", Name = "tot", Est = mcbs_tot_est$est, Var = mcbs_tot_est$var)

@@

    if (entropy == "EL") {
      w <- -d2i / drop(zmat %*% lam)
    } else if (entropy == "ET") {
      w <- d2i * exp(drop(zmat %*% lam))
    } else {
      stop("We only accept entropy of EL or ET.")
    }

    if (returnw) {
      return(w)
    } else {
      return(drop((w * w1i * qi) %*% zmat) - T1)
    }
  }

  j <- function(lam, zmat, w1i, d2i, T1, qi, entropy) {
    if (entropy == "EL") {
      res <- t(zmat) %*% diag((w1i * qi * d2i) / drop(zmat %*% lam)^2) %*% zmat
    } else if (entropy == "ET") {
      res <- t(zmat) %*% diag((w1i * qi * d2i) * exp(drop(zmat %*% lam))) %*% zmat
    }

@@

  if (!(res$termcd %in% c(1))) {
    return(NA)
  }

@@

data_int_mcbsacs <- function(mcbs_all, cov_ests) {

  qi = 1
  entropy = "EL"

  # Default: p0_df -> mcbs_all
  # d_vec <- 1 / p0_df$pi0
  d_vec <- mcbs_all$PUFSWGT
  if (entropy == "EL") {
    gdi <- -1 / d_vec
    # FIXME: We need an estimate of gdip
    gdip <- -1 / (1 / pop_df$pi0)
    # END FIXME:
    gpinv <- d_vec^2
  } else if (entropy == "ET") {
    gdi <- log(d_vec)
    gdip <- log(1 / pop_df$pi0)
    gpinv <- d_vec
  } else {
    stop("This type of entropy has not yet been implemented.")
  }

  samp_X <- model.matrix(~edulths + edueqhs + 
                         marmar + marwid + mardiv + 
                         sexm + racew + raceb + raceh, data = mcbs_all)
  zmat <- cbind(samp_X, gdi)
  w1i <- 1

  c_ests <- filter(cov_ests, !(Name %in% c("edumths", "sexf", "marsin")))
  xhat <- c_ests$Est
  # Columns: tot, edulths, edueqhs, marmar, marwid, mardiv, sexm,
  #          racew, raceb, raceh
  # Rows: same dimension as length(xhat)
  x_mat <- matrix(c(as.numeric(c_ests$Name == "tot"),
                    as.numeric(c_ests$Name == "edulths"),
                    as.numeric(c_ests$Name == "edueqhs"),
                    as.numeric(c_ests$Name == "marmar"),
                    as.numeric(c_ests$Name == "marwid"),
                    as.numeric(c_ests$Name == "mardiv"),
                    as.numeric(c_ests$Name == "sexm"),
                    as.numeric(c_ests$Name == "racew"),
                    as.numeric(c_ests$Name == "raceb"),
                    as.numeric(c_ests$Name == "raceh")),
                  ncol = 10)

  vmat <- diag(c_ests$Var)
  vinv <- diag(1 / c_ests$Var)
  xgls <- drop(solve(t(x_mat) %*% vinv %*% x_mat) %*% 
              (t(x_mat) %*% vinv %*% xhat))
  T1 <- c(xgls, sum(gdip * qi))
  dc_w <- solveGE(zmat, w1i, d_vec, T1, qi, entropy)

  # Mean Estimation
  N_pop <- filter(cov_ests, Data == "CMS", Name = "tot") %>% pull(Est)
  Y_vec <- as.numeric(mcbs_all$CVS_ONEDOSE == "1")
  theta <- sum(Y_vec * dc_w * w1i) / N_pop

  z_dc <- sampX
  gam_dc <- 
    solve(t(z_dc) %*% diag(d_vec * gpinv * qi) %*% z_dc,
          t(z_dc) %*% diag(d_vec * gpinv) %*% Y_vec)
  coefs <- as.numeric(gam_dc)
  u_mat <- sampX
  eps <- (Y_vec - drop(u_mat %*% coefs))

@@

@@

Next steps:
* Use data integration model.

>>>>

library(dplyr)
library(readr)
library(haven)
library(stringr)
library(purrr)
library(nleqslv)
library(doParallel)

@@

```{r}
#| label: acs uncertainty

acs_cl <- read_csv("data/cleaned_acs_data.csv")
acs_med <- filter(acs_cl, HINS3 == 1)

@@

mcbs_all <- mcbs_all %>%
  mutate(CVS_ONEDOSE = case_when(
    CVS_ONEDOSE == "1" ~ 1,
    CVS_ONEDOSE == "2" ~ 2,
    is.na(CVS_ONEDOSE) & (CVS_TWODOSE %in% c("1", "2")) ~ as.numeric(CVS_TWODOSE),
    CVS_ONEDOSE %in% c("R", "D") & !is.na(CVS_VCNUMS) ~ 1,
    TRUE ~ 2
  ))

@@

```{r}
#| label: add CMS data

cov_ests <- 
  # bind_rows(
  purrr::map_dfr(param_df$varname, function(vn) {
    out <- get_est_uncertainty(acs_med, vn)
    return(tibble(Data = "ACS", Name = vn, Est = out$est, Var = out$var))
  })#,
  # purrr::map_dfr(param_df$varname[-1], function(vn) {
  #   out <- get_est_uncertainty(mcbs_all, vn, def_w = "PUFSWGT")
  #   return(tibble(Data = "MCBS", Name = vn, Est = out$est, Var = out$var))
  # }))

@@

cov_ests <- cov_ests %>% 
  add_row(Data = "CMS", Name = "tot", Est = 63892626, Var = 1) %>%
  add_row(Data = "CMS", Name = "sexm", Est = 29159084, Var = 1) %>%
  add_row(Data = "CMS", Name = "sexf", Est = 34733542, Var = 1) %>%
  mutate(Name = ifelse(Name %in% c("HINS3", "ones"), "tot", Name))

# %>%
#  add_row(Data = "MCBS", Name = "tot", Est = mcbs_tot_est$est, Var = mcbs_tot_est$var)

@@

    if (entropy == "EL") {
      w <- (-d2i) / drop(zmat %*% lam)
    } else if (entropy == "ET") {
      w <- d2i * exp(drop(zmat %*% lam))
    } else {
      stop("We only accept entropy of EL or ET.")
    }

    if (returnw) {
      return(w)
    } else {
      return(T1 - drop((w * w1i * qi) %*% zmat))
    }
  }

  j <- function(lam, zmat, w1i, d2i, T1, qi, entropy) {
    if (entropy == "EL") {
      res <- (-1) * t(zmat) %*% diag((w1i * qi * d2i) / drop(zmat %*% lam)^2) %*% zmat
    } else if (entropy == "ET") {
      res <- (-1) * t(zmat) %*% 
        diag((w1i * qi * d2i) * exp(drop(zmat %*% lam))) %*% zmat
    }

@@

  if (!(res$termcd %in% c(1, 2))) {
    return(NA)
  }

@@

# The solveGE_alpha function estimates alpha and the weights.
# This is useful for when we do not have design weights for the entire
# population.
solveGE_alpha <- function(alpha, zmat, w1i, d2i, T1, qi, entropy, retw = FALSE) {

  N <- T1[1]
  T1_adj <- T1
  T1_adj[length(T1_adj)] <- alpha * N
  dc_w <- solveGE(zmat, w1i, d2i, T1_adj, qi, entropy)

  if (retw) {
    return(dc_w)
  }

  Ka_type <- "alpha"
  if (Ka_type == "alpha") {
    # HACK: G(x) = -log(x) for EL only
    G <- function(x) {-log(x)}
    return(sum(G(dc_w)) - N * alpha)
  }
}

data_int_mcbsacs <- function(mcbs_all, cov_ests, weight_v = "PUFSWGT") {

  qi = 1
  entropy = "EL"
  N_pop <- filter(cov_ests, Data == "CMS", Name == "tot") %>% pull(Est)

  # Default: p0_df -> mcbs_all
  # d_vec <- 1 / p0_df$pi0
  d_vec <- mcbs_all[[weight_v]]
  if (entropy == "EL") {
    gdi <- -1 / d_vec
    gpinv <- d_vec^2
  } else if (entropy == "ET") {
    gdi <- log(d_vec)
    gpinv <- d_vec
  } else {
    stop("This type of entropy has not yet been implemented.")
  }

  samp_X <- model.matrix(~edulths + edueqhs + marmar + marwid + mardiv + sexm,
                        data = mcbs_all)
  zmat <- cbind(samp_X, gdi)
  w1i <- 1

  c_ests <- 
    filter(cov_ests, !(Name %in% c("edumths", "sexf", "marsin"))) %>%
    filter(!(Name %in% c("racew", "raceb", "raceh")))

  # We want the mean for X not the total
  xhat <- c_ests$Est / N_pop
  # Columns: tot, edulths, edueqhs, marmar, marwid, mardiv, sexm,
  #          racew, raceb, raceh
  # Rows: same dimension as length(xhat)
  x_mat <- matrix(c(as.numeric(c_ests$Name == "tot"),
                    as.numeric(c_ests$Name == "edulths"),
                    as.numeric(c_ests$Name == "edueqhs"),
                    as.numeric(c_ests$Name == "marmar"),
                    as.numeric(c_ests$Name == "marwid"),
                    as.numeric(c_ests$Name == "mardiv"),
                    as.numeric(c_ests$Name == "sexm")),
                  ncol = length(unique(c_ests$Name)))

  # We want the variance of the mean not the variance of the total
  vmat <- diag(c_ests$Var / N_pop^2)
  vinv <- diag(N_pop^2 / c_ests$Var)
  xgls <- drop(solve(t(x_mat) %*% vinv %*% x_mat) %*% 
              (t(x_mat) %*% vinv %*% xhat))

  # We want T1 to be the totals
  T1 <- c(xgls * N_pop, NA)

  # Estimating Alpha
  alpha_init <- (-1) * nrow(mcbs_all) / N_pop
  a_res <- 
    optim(par = alpha_init, fn = solveGE_alpha, 
        zmat = zmat, w1i = w1i, d2i = d_vec, T1 = T1, qi = qi, entropy = entropy)

  if (a_res$convergence != 0) {
    warning(paste0("a_res has convergence", a_res$convergence))
  }

  dc_w <- 
    solveGE_alpha(a_res$par, zmat, w1i, d_vec, T1, qi, entropy, retw = TRUE)

  # Mean Estimation
  Y_vec <- as.numeric(mcbs_all$CVS_ONEDOSE == 1)
  theta <- sum(Y_vec * dc_w * w1i) / N_pop

  # FIXME: Do we want samp_X or zmat?
  gam_dc <- 
    solve(t(samp_X) %*% diag(d_vec * gpinv * qi) %*% samp_X,
          t(samp_X) %*% diag(d_vec * gpinv) %*% Y_vec)
  coefs <- as.numeric(gam_dc)
  eps <- (Y_vec - drop(samp_X %*% coefs))

@@

reg_est <- function(mcbs_all, cov_ests, weight_v = "PUFSWGT") {

  N_pop <- filter(cov_ests, Data == "CMS", Name == "tot") %>% pull(Est)
  Y_vec <- as.numeric(mcbs_all$CVS_ONEDOSE == 1)
  mod_df <- mcbs_all %>%
    mutate(Y = (CVS_ONEDOSE == 1))

  mod <- lm(Y ~ sexm, data = mod_df, weights = mod_df[[weight_v]])
  yhat <- predict(mod, mod_df)

  # TEST:
  # mod_t <- lm(Y ~ edulths + edueqhs + marmar + marwid + mardiv + sexm,
  #   data = mod_df, weights = mod_df[[weight_v]])
  # yhat <- predict(mod_t, mod_df) 
  # END TEST:

  # Regression Estimation
  cms_tot <- filter(cov_ests, Data == "CMS", Name == "tot") %>% pull(Est)
  cms_sexm <- filter(cov_ests, Data == "CMS", Name == "sexm") %>% pull(Est)
  cms_mean <- cms_tot / N_pop
  cms_meansexm <- cms_sexm / N_pop

  err_term <- sum((mod_df$Y - yhat) * mod_df[[weight_v]]) / N_pop

  theta <- 
    cms_mean * mod$coefficients[1] + cms_meansexm * mod$coefficients[2] + err_term

  # Variance Estimation
  pihat <- 1 / mod_df[[weight_v]]
  v_hat <- sum((1 - pihat) * (mod_df$Y - yhat)^2 * (1 / pihat)^2) / N_pop^2

  return(list(theta = theta, v_hat = v_hat))
}

ht_est <- function(mcbs_all, cov_ests, weight_v = "PUFSWGT") {

  N_pop <- filter(cov_ests, Data == "CMS", Name == "tot") %>% pull(Est)
  Y_vec <- as.numeric(mcbs_all$CVS_ONEDOSE == 1)
  theta <- sum(Y_vec * mcbs_all[[weight_v]]) / N_pop
  pihat <- 1 / mod_df[[weight_v]]
  v_hat <- sum((1 - pihat) * (Y_vec  / pihat)^2) / N_pop^2

  return(list(theta = theta, v_hat = v_hat))
}

@@

# Assessing the Variance

```{r}
#| label: variance analysis of mcbs

res <- data_int_mcbsacs(mcbs_all, cov_ests, weight_v = "PUFSWGT") 
ms <- reg_est(mcbs_all, cov_ests)
ht <- ht_est(mcbs_all, cov_ests)

cl <- makeCluster(min(detectCores() - 1, 100), outfile = "")
registerDoParallel(cl)

packs <- c("dplyr", "stringr", "purrr", "nleqslv")

mc_res <- 
  foreach(b = 1:100, .packages = packs,
          .errorhandling = "remove", .inorder = FALSE) %dopar% {

  if ((b %% 10) == 0) {
    print(paste0("Iteration: ", b))
  }

  weight_v <- paste0("PUFS", str_sub(paste0("00", b), -3))
  ms <- data_int_mcbsacs(mcbs_all, cov_ests, weight_v = weight_v) 
  # ms <- reg_est(mcbs_all, cov_ests, weight_v = weight_v)

  return(
      tibble(Est = c("MSEst"),
             Theta = c(ms$theta),
             Var = c(ms$v_hat), 
             Iter = b)
  )

  } %>% bind_rows()

stopCluster(cl)

# Weighted variance estimate
var(mc_res$Theta)

# Estimated variance estimate
res$v_hat

# Relative Bias
(res$v_hat - var(mc_res$Theta)) / var(mc_res$Theta)

Next steps:
* Use data integration model.
* Check variance assessment using other weight models (PUFS001, etc.).

### 2024-11-12 src/explore/20241101-msdcsim_estalpha.qmd
Finish all of the simulations. Added simulation script.

  j <- function(lam, zmat, w1i, d2i, T1, qi, entropy) {
    end <- ncol(zmat)
    if (entropy == "EL") {
      res <- (-1) * t(zmat) %*% diag((w1i * qi * d2i) / drop(zmat %*% lam)^2) %*% zmat
    } else if (entropy == "ET") {
      res <- (-1) * t(zmat) %*% 
        diag((w1i * qi * d2i) * exp(drop(zmat %*% lam))) %*% zmat
    }

@@

  # Estimating Alpha
  alpha_init <- (-1) * nrow(p0_df) / N_obs
  a_res <- 
    nlm(f = solveGE_alpha, p = alpha_init,
        zmat = zmat, w1i = w1i, d2i = d_vec, T1 = T1, qi = qi, entropy = entropy)

  if (a_res$code != 1) {
    warning(paste0("a_res has code", a_res$code))
  }

  dc_w <- 
    solveGE_alpha(a_res$estimate, zmat, w1i, d_vec, T1, qi, entropy, retw = TRUE)

@@

mc_res <- 
  foreach(iter=1:B_sims,
          .packages = c("dplyr", "nleqslv"),
          .options.RNG = 1) %dorng% {

    if (iter %% 100 == 0) {
      print(paste0("Iter: ", iter))
    }

@@

    ht <- ht_est(p0_df, N_obs)
    reg <- reg_est(p0_df, p1_df, N_obs)
    ms_est <- dc_ms_ybar(p0_df, p1_df, p2_df, pop_df)

@@

# Analyze Mean
mc_res %>%
  mutate(err = Theta - true_theta) %>%
  mutate(CI = abs(err) < qnorm(0.975) * sqrt(Var)) %>%
  group_by(Est) %>%
  summarize(Bias = mean(err, na.rm = TRUE),
            SE = sd(err, na.rm = TRUE),
            RMSE = sqrt(mean(err^2, na.rm = TRUE)),
            EmpCI = mean(CI, na.rm = TRUE),
            MCVar = var(Theta, na.rm = TRUE),
            EstVar = mean(Var, na.rm = TRUE),
            sdest = sd(Theta, na.rm = TRUE),
            RelBias = (EstVar - MCVar) / MCVar) %>%
  ungroup() %>%
  mutate(Ttest = abs(Bias) / sqrt(sdest^2 / B_sims)) %>%
  select(-sdest) %>%
  mutate(Ind = case_when(
    Est == "MSReg" ~ 5,
    Est == "MSEst" ~ 4,
    Est == "MSPop" ~ 3,
    Est == "NNReg" ~ 2,
    Est == "HT" ~ 1)
  )

>>>>

  j <- function(lam, zmat, w1i, d2i, T1, qi, entropy) {
    if (entropy == "EL") {
      res <- (-1) * t(zmat) %*% diag((w1i * qi * d2i) / drop(zmat %*% lam)^2) %*% zmat
    } else if (entropy == "ET") {
      res <- (-1) * t(zmat) %*% 
        diag((w1i * qi * d2i) * exp(drop(zmat %*% lam))) %*% zmat
    }

@@

  # Estimating Alpha
  alpha_init <- (-1) * nrow(p0_df) / N_obs
  a_res <- 
    optim(par = alpha_init, fn = solveGE_alpha, 
        zmat = zmat, w1i = w1i, d2i = d_vec, T1 = T1, qi = qi, entropy = entropy)

  if (a_res$convergence != 0) {
    warning(paste0("a_res has convergence", a_res$convergence))
  }

  dc_w <- 
    solveGE_alpha(a_res$par, zmat, w1i, d_vec, T1, qi, entropy, retw = TRUE)

@@

mc_res <- 
  foreach(iter=1:B_sims, .packages = c("dplyr", "nleqslv"), .options.RNG = 1) %dopar% {

    if (iter %% 50 == 0) {
      print(paste0("Iter: ", iter))
      # set.seed(iter)
    }

@@

    ht <- ht_est(p0_df, N_obs)
    reg <- reg_est(p0_df, p1_df, N_obs)
    ms_est <- suppressWarnings(dc_ms_ybar(p0_df, p1_df, p2_df, pop_df))

@@

# Analyze Mean
mc_res %>%
  mutate(err = Theta - true_theta) %>%
  mutate(CI = abs(err) < qnorm(0.975) * sqrt(Var)) %>%
  group_by(Est) %>%
  summarize(Bias = mean(err, na.rm = TRUE),
            SE = sd(err, na.rm = TRUE),
            RMSE = sqrt(mean(err^2, na.rm = TRUE)),
            EmpCI = mean(CI, na.rm = TRUE),
            MCVar = var(Theta, na.rm = TRUE),
            EstVar = mean(Var, na.rm = TRUE),
            sdest = sd(Theta, na.rm = TRUE),
            RelBias = (EstVar - MCVar) / MCVar) %>%
  ungroup() %>%
  mutate(Ttest = abs(Bias) / sqrt(sdest^2 / B_sims)) %>%
  select(-sdest)
